A comparison of classification algorithms applied to the Adult Database

### 2009-12-10 report/final.tex
added ILP experimental description.

\documentclass[a4paper]{llncs}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage[utf8]{inputenc}
\usepackage{subfigure}
\usepackage{a4wide}
\usepackage{amsmath}
\usepackage{listings}
\usepackage[hyphens]{url}
\urldef{\mails}\path|{ei05011,ei05028}@fe.up.pt|
\bibliographystyle{splncs}
\begin{document}
\title{A comparison of classification algorithms applied to the Adult Database}
\author{Flávio Cruz \and João Azevedo}
\institute{Faculdade de Engenharia da Universidade do Porto\\
    Rua Dr. Roberto Frias, s/n 4200-465 Porto PORTUGAL\\
    \mails}
\maketitle

@@

@@

\clearpage
\section{Classification Rules}
\begin{lstlisting}[language=c,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Subset of the ordered set of rules induced by CN2 with the laplacian error estimation method},label=cn2_rules]
IF    age < 20.50
  AND capital_gain < 9720.00
  AND hours_per_week < 59.00
THEN  class = less_or_equal_than_50K  [0 2366]
ELSE
IF    19.50 < age < 21.50
  AND marital_status = never_married
  AND capital_gain < 67047.00
THEN  class = less_or_equal_than_50K  [0 664]
ELSE
IF    age < 54.50
  AND capital_gain > 10585.50
  AND hours_per_week > 33.00
THEN  class = more_than_50K  [520 0]
ELSE
IF    age < 23.50
  AND hours_per_week < 31.00
THEN  class = less_or_equal_than_50K  [0 492]
ELSE
IF    age < 24.50
  AND workclass = private
  AND relationship = own_child
  AND capital_loss < 2129.50
  AND hours_per_week > 32.50
  AND native_country = united_states
THEN  class = less_or_equal_than_50K  [0 562]
ELSE
IF    age < 60.50
  AND 7565.50 < capital_gain < 30961.50
THEN  class = more_than_50K  [445 0]
ELSE
IF    relationship = own_child
  AND race = black
  AND hours_per_week < 49.00
THEN  class = less_or_equal_than_50K  [0 261]
ELSE
IF    age < 38.50
  AND education = hs_grad
  AND marital_status = never_married
  AND relationship = own_child
  AND hours_per_week < 75.00
THEN  class = less_or_equal_than_50K  [0 378]
ELSE
IF    20.50 < age < 23.50
  AND workclass = private
  AND marital_status = never_married
  AND capital_loss < 2116.00
  AND hours_per_week > 32.50
THEN  class = less_or_equal_than_50K  [0 434]
\end{lstlisting}

>>>>

\documentclass[a4paper]{llncs}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage[utf8]{inputenc}
\usepackage{subfigure}
\usepackage{a4wide}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{multirow}
\usepackage[hyphens]{url}
\urldef{\mails}\path|{ei05011,ei05028}@fe.up.pt|
\bibliographystyle{splncs}
\begin{document}
\title{A comparison of classification algorithms applied to the Adult Database}
\author{Flávio Cruz \and João Azevedo}
\institute{Faculdade de Engenharia da Universidade do Porto\\
    Rua Dr. Roberto Frias, s/n 4200-465 Porto PORTUGAL\\
    \mails}
\maketitle

@@

\subsection{ILP}

There are a lot of techniques for building ILP classifiers. Due to time
constraints (as building an ILP classifier is usually a very time-consuming 
task), we only tested the basic Aleph system, which uses inverse entailment for
building the bottom clause and a best-first search method to find the best
hypothesis. The results obtained are described in table \ref{tbl:results_ilp}.

\begin{table}
\begin{center}
\begin{tabular}{ | c | c | c | c | c |}
\hline
\vspace{1pt} & \multicolumn{4}{| c |}{\textbf{Actual}} \\ \hline
\multirow{3}{*}{\textbf{Predicted}} & \vspace{1pt} & \textbf{+} & \textbf{-} & \textbf{Total} \\ \hline
& \textbf{+} & 647 & 133 & 780 \\ \hline
& \textbf{-} & 3199 & 12302 & 15501 \\ \hline
& \textbf{Total} & 3846 & 12435 & 16281 \\ \hline
\end{tabular}
\caption{Results for ILP (Aleph).}
\label{tbl:results_ilp}
\end{center}
\end{table}

The error rate produced by Aleph when evaluating the test data was 20.45\%.
18881882 clauses were built, taking around 5 hours for the whole training
data. Despite the number of clauses, most of them are irrelevant, directly
classifying examples where no more general rule is available. Listing 
\ref{ilp_clauses} lists the most relevant clauses produced.

@@

\clearpage
\section{Classification Rules}
\begin{lstlisting}[language=c,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Subset of the ordered set of rules induced by CN2 with the laplacian error estimation method},label=cn2_rules]
IF    age < 20.50
  AND capital_gain < 9720.00
  AND hours_per_week < 59.00
THEN  class = less_or_equal_than_50K  [0 2366]
ELSE
IF    19.50 < age < 21.50
  AND marital_status = never_married
  AND capital_gain < 67047.00
THEN  class = less_or_equal_than_50K  [0 664]
ELSE
IF    age < 54.50
  AND capital_gain > 10585.50
  AND hours_per_week > 33.00
THEN  class = more_than_50K  [520 0]
ELSE
IF    age < 23.50
  AND hours_per_week < 31.00
THEN  class = less_or_equal_than_50K  [0 492]
ELSE
IF    age < 24.50
  AND workclass = private
  AND relationship = own_child
  AND capital_loss < 2129.50
  AND hours_per_week > 32.50
  AND native_country = united_states
THEN  class = less_or_equal_than_50K  [0 562]
ELSE
IF    age < 60.50
  AND 7565.50 < capital_gain < 30961.50
THEN  class = more_than_50K  [445 0]
ELSE
IF    relationship = own_child
  AND race = black
  AND hours_per_week < 49.00
THEN  class = less_or_equal_than_50K  [0 261]
ELSE
IF    age < 38.50
  AND education = hs_grad
  AND marital_status = never_married
  AND relationship = own_child
  AND hours_per_week < 75.00
THEN  class = less_or_equal_than_50K  [0 378]
ELSE
IF    20.50 < age < 23.50
  AND workclass = private
  AND marital_status = never_married
  AND capital_loss < 2116.00
  AND hours_per_week > 32.50
THEN  class = less_or_equal_than_50K  [0 434]
\end{lstlisting}
\clearpage

\begin{lstlisting}[language=c,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Subset of the prolog clauses produced by Aleph},label=ilp_clauses]
more_than_50K(A) :-
   middle_aged(A), very_high_capital_gain(A), is_native_of_united_states(A).
more_than_50K(A) :-
   education_prof_school(A), wife(A), workaholic(A).
more_than_50K(A) :-
   senior(A), work_private(A), very_high_capital_gain(A).
more_than_50K(A) :-
   senior(A), work_federal_gov(A), education_doctorate(A).
more_than_50K(A) :-
   education_doctorate(A), medium_capital_loss(A), workaholic(A).
more_than_50K(A) :-
   work_self_emp_inc(A), high_capital_gain(A), full_time_worker(A).
more_than_50K(A) :-
   middle_aged(A), education_masters(A), is_native_of_iran(A).
more_than_50K(A) :-
   senior(A), very_high_capital_gain(A), over_time_worker(A).
more_than_50K(A) :-
   other_relative(A), high_capital_gain(A).
more_than_50K(A) :-
   education_bachelors(A), occupation_other_service(A), high_capital_gain(A).
more_than_50K(A) :-
   senior(A), married_civ_spouse(A), very_high_capital_gain(A).
more_than_50K(A) :-
   senior(A), full_time_worker(A), is_native_of_cambodia(A).
more_than_50K(A) :-
   middle_aged(A), married_civ_spouse(A), is_native_of_thailand(A).
more_than_50K(A) :-
   young(A), high_capital_gain(A), over_time_worker(A).
more_than_50K(A) :-
   middle_aged(A), asian_pac_islander(A), high_capital_gain(A).
more_than_50K(A) :-
   occupation_protective_serv(A), high_capital_gain(A), full_time_worker(A).
more_than_50K(A) :-
   work_state_gov(A), occupation_exec_managerial(A), high_capital_gain(A).
more_than_50K(A) :-
   education_bachelors(A), occupation_tech_support(A), wife(A).
more_than_50K(A) :-
   education_prof_school(A), very_high_capital_gain(A).
more_than_50K(A) :-
   high_capital_gain(A), full_time_worker(A), is_native_of_canada(A).
more_than_50K(A) :-
   education_prof_school(A), medium_capital_loss(A), over_time_worker(A).
more_than_50K(A) :-
   work_private(A), education_doctorate(A), is_native_of_taiwan(A).
\end{lstlisting}
